\documentclass{article}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{hhline}

\title{SLURM: Simple Linux Utility for Resource Management\thanks{
This document was prepared as an account of work sponsored by an
agency of the United States Government.  Neither the United States
Government nor the University of California nor any of their
employees, makes any warranty, express or implied, or assumes any
legal liability or responsibility for the accuracy, completeness, or
usefulness of any information, apparatus, product, or process
disclosed, or represents that its use would not infringe privately
owned rights. Reference herein to any specific commercial product,
process, or service by trade name, trademark, manufacturer, or
otherwise, does not necessarily constitute or imply its endorsement,
recommendation, or favoring by the United States Government or the
University of California.  The views and opinions of authors expressed
herein do not necessarily state or reflect those of the United States
Government or the University of California, and shall not be used for
advertising or product endorsement purposes.
This work was performed under the auspices of the U. S. Department of
Energy by the University of California, Lawrence Livermore National
Laboratory under Contract No. W-7405-Eng-48.}}

\author{Morris Jette \and Mark Grondona}

% We cheat here to easily get the desired allignment 
\date{\{jette1,mgrondona\}@llnl.gov}

\begin{document}

\maketitle

\begin{abstract}
Simple Linux Utility for Resource Management (SLURM) is an open source,
fault-tolerant, and highly scalable cluster management and job 
scheduling system for Linux clusters of thousands of nodes.  Components 
include machine status, partition management, job management, scheduling 
and stream copy modules.  This paper presents a overview of the SLURM architecture and functionality.
\end{abstract}

\section{Overview}

TBD

\end{document}
